---
title: "Practical Machine Learning Assignment"
author: "Firdaus Afifi"
date: "May 11, 2017"
output: html_document
---

## Introduction

One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, our goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.who were asked to perform barbell lifts correctly and incorrectly in 5 different ways.`Random Forest` is apply it to a test dataset to make predictions. This machine learning model shows accuracy of 99% on a training dataset.

##Data

The data for this assignment come in the form of csv file.
It was taken from the Human Activity Recognition programme at [Groupware](http://groupware.les.inf.puc-rio.br/har) and its information can be found under the section Weight Lifting Exercise Dataset.

First we download the file from the links below.

```{r}
training.data.url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test.data.url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(training.data.url, "pml-training.csv")
download.file(test.data.url, "pml-test.csv")
```

Then we perform data importing and cleansing.

```{r}
read.data <- function(data.name) {
  # import data
  file.name <- paste("pml-", data.name, ".csv", sep="")
  data <- read.csv(file.name, header=TRUE, na.strings=c("", "NA", "#DIV/0!"))
  # remove variables with NAs or empty
  data <- data[, -which(sapply(data, function(x) any(is.na(x)) || any(x=="")))]
  # remove unnecessary variables
  rm.cols <- c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")
  data <- data[, -which(names(data) %in% rm.cols)]
  data
}
training.data <- read.data("training")
test.data <- read.data("test")
```

Next we split the training dataset intro training and validation data, with the rate 60:40 respectively.

```{r}
library(caret)
trainingIndex <- createDataPartition(training.data$classe, p=.60, list=FALSE)
training.train <- training.data[trainingIndex, ]
training.validation <- training.data[-trainingIndex, ]
```

##Machine Learning Modeling

We use `Random Forest` method for our machine learning model with 5-fold cross validation.

```{r}
library(randomForest)
tc <- trainControl(method = "cv", number = 5)
rf.fit <- train(classe ~ ., data=training.train, method="rf", trControl=tc)
```

##Results

We now compare the results from the predition with the actual data.

```{r}
classe.idx <- which(names(training.validation)=="classe")
confusionMatrix(predict(rf.fit, newdata=training.validation[, -classe.idx]),training.validation$classe)
```

